---
title: "In_class exercise 10.9"
author: "Feiyang CHen"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
d <- read_csv("/Users/frankchan/Library/CloudStorage/OneDrive-UniversityofGlasgow/github/MY_POLI3148_23Fall/_DataPublic_/vdem/1984_2022/vdem_1984_2022_external.csv")
```

In-Class Exercise

The quality of education has a decisive effect on a country's future development. Applying the data wrangling tools we introduce in this lecture, perform the following task:

1.  **Codebook lookup**. Look up the codebook, answer the following questions:
    1.  What indicators regarding the quality of education are available in the V-Dem datasets?

        1.  Education 15+ (E) (e_peaveduc)

            Educational inequality, Gini (E) (e_peedgini)

    2.  What are the data's coverage (i.e., for which countries and years do we have data?)

        1900-2022

    3.  What are their sources? Provide the link to least 1 source.

        Pemstein et al. (2023, V-Dem Working Paper Series 2023:21); V-Dem Codebook (see suggested citation at the top of this document)
2.  **Subset by columns**
    1.  Create a dataset containing only the country-year identifiers and indicators of education quality.

    2.  Rename the columns of education quality to make them informative.
3.  **Subset by rows**
    1.  List 5 countries-years that have the highest education level among its population.

    2.  List 5 countries-years that suffer from the most severe inequality in education.
4.  **Summarize the data**
    1.  Check data availability: For which countries and years are the indicators of education quality available?

    2.  Create two types of country-level indicators of education quality

        1.  Average level of education quality from 1984 to 2022

        2.  Change of education quality from 1984 to 2022

    3.  Examine the data and *briefly* discuss: Which countries perform the best and the worst in terms of education quality in the past four decades?

Submission **requirement:** You will submit your outputs through Moodle. In your submission:

1.  Attach a PDF document rendered by Rmarkdown
2.  In the text field of your submission, include the link to the corresponding Rmarkdown file in your *DaSPPA portfolio* GitHub repo.

1.1 Education 15+ (E) (e_peaveduc), Educational inequality, Gini (E) (e_peedgini)

1.2 1900-2022, all countries

```{r}
d_education <- d |> 
  select(country_name, year, e_peaveduc, e_peedgini)

str(d_education)

```

1.3 Pemstein et al. (2023, V-Dem Working Paper Series 2023:21); V-Dem Codebook (see suggested citation at the top of this document)

2.1

```{r}
d_education <- d |> 
  select(country_name, year, e_peaveduc, e_peedgini)

d_education
```

2.2

```{r}
d_education <- d_education |>
  rename("Education_level" = "e_peaveduc") |>
  rename("Educational_inequality" = "e_peedgini") |>
  rename("Country" = "country_name") |>
  rename("Year" = "year")

```

3.1

```{r}
d_education |> 
  slice_max(order_by = Education_level, n = 5)
```

3.2

```{r}
d_education |> 
  slice_max(order_by = Educational_inequality, n = 5)
```

4.1

```{r}
d_education_filter <- d_education[!is.na(d_education$Educational_inequality), ]
d_education_filter 

d_education_filter |> 
  select(Country, Year) 
```

Create two types of country-level indicators of education quality

-   Average level of education quality from 1984 to 2022

-   Change of education quality from 1984 to 2022

4.2.1

```{r}
d_education_filter <- d_education[!is.na(d_education$Educational_inequality), ]

selected <- d_education_filter |> 
  select(Country, Education_level) 
```

```{r}
averages <- selected |> group_by(Country) |> summarise(avg_el = mean(Education_level))
averages
```

4.2.2

```{r}
d_education_filter_2 <- d_education[!is.na(d_education$Education_level), ]

change <- d_education_filter_2 |> 
  select(Country, Year, Education_level) |> 
  group_by(Country) |> 
  summarise(change = Education_level[which.max(Year)] - Education_level[which.min(Year)])

change
```

4.3

Examine the data and *briefly* discuss: Which countries perform the best and the worst in terms of education quality in the past four decades?
